Effects of Tunable Hydrophobicity on the Collective Hydrodynamics of Janus Particles under Flows

### 2023-03-21 doc/PRF2022/PRF_main.tex
finished up small changes. Now need to fix Fig 6 and area

\begin{abstract}
  Active colloidal systems with non-equilibrium self-organization is a
  long-standing, challenging area in material sciences and biology. To
  understand how hydrodynamic flow may be used to actively control
  self-assembly of Janus particles (JPs), we developed a model for the
  many-body hydrodynamics of amphiphilic JPs suspended in a viscous
  fluid with imposed far-field background flows (Journal of Fluid
  Mechanics, 941, 2022). In this work we alter the hydrophobic
  distribution on the JP-solvent interface to investigate the
  hydrodynamics that underlies the various morphologies and rheological
  properties of the JP assembly in the suspension. We find JPs assemble
  into uni-lamella, multi-lamella, and striated structures. To introduce
  dynamics, we include a planar linear shear flow and a steady
  Taylor-Green mixing flow, and measure the collective dynamics of JP
  particles in terms of their (a) free energy from the hydrophobic
  interactions between the JPs, (b) order parameter for the ordering of
  JPs in terms of alignment of their directors, and (c) strain parameter
  that captures the deformation in the assembly. We characterize the
  effective material properties of the JP structures and find that the
  uni-lamellar structures increases orientation order under shear flow,
  the multilamellar structure behaves as a shear thinning fluid, and the
  striated structure possesses a yield stress. These numerical results
  provide insights into dynamic control of non-equilibrium active
  biological systems with similar self-organization.
\end{abstract}

@@

The shear stress-strain rate data are excellently fit by power laws.
Figures~\ref{fig:rheologysummary}(a)--(c) show the data for Type I, Type
II and Type III JPs, respectively, under both shear flow (squares) and
TG flow (triangles). The bilayer phases consisting of Type I JPs are
basically Newtonian with viscosity $0.014$ MPa\;$\upmu$s $= 14$ mPa\;s,
about 14 times the viscosity of water at room temperature
(Figures~\ref{fig:rheologysummary}(a), dashed curve). The vesicle
bilayer phase is slightly shear thickening
(Figure~\ref{fig:rheologysummary}(a), solid curve). The multilamellar
phase consisting of Type II JPs is somewhat shear thinning, but its
viscosity is larger than for Type I. Finally, the striated phase is
strongly shear-thinning. This further explains the rigid-like behavior
of the striated phase found in the strain, shear-rate relationships from
Figures~\ref{fig:s2esummary}(a) and (b) (triangles).
\todo[inline]{RBC flow refereces}

@@

In this work, we employ the newly developed JP model using
BIEs~\cite{Fu20, Fu2022_JFM} and tune the boundary conditions with
energy normalization to study the collective dynamics of amphiphilic (BC
I), biased hydrophobic (BC II), and bipolar (BC III) JPs under various
flowing conditions (quiescent flow, linear shear flow, and Taylor-Green
flow). Three quantities are computed to characterize the dynamics of the
collective configurations of JPs: free energy $F$, strain parameter $E$,
and scalar order parameter $S_2$. 
%We quantify how
%the collective hydrodynamics of JPs is affected by an external flow.
Under a given flow, we use these three measures of deformations to
quantify the differences in the collective dynamics between the three
types of JPs. These results, summarized below, provide general insight
into the dynamic control of active particles in a viscous suspension.

Within the framework of boundary integral equations, we derive a
convenient expression for computing the free energy using the single
layer potential (see equation~\eqref{eq:normal_deriv}). In a quiescent
flow, the free energy profiles demonstrate that the relaxation process
for particles confined in a certain size of box is independent of the
number of particles. However, we find that the final configurations does
depend on the initial distribution of particle directors $\dd_i$.
Therefore, multiple patterns or local energy minimum states may appear
depending on the initial setup.

Under a relatively weak linear shear flow, the amphiphilic JPs behave as
a unilamellar vesicle that elongates and tank-treads, with the scalar
order parameter increasing over time. The assembly of multi-lamella and
striated JP structures resemble a rigid body motion with minimal
deformation. The effective viscosity of the material quantitatively
validates this result. High shear-rate cases provide a range of critical
shear rates where the structures break apart and undergo topological
changes. We further show that free energy, scalar order parameter, and
strain are effective measures to quantitatively capture the collective hydrodynamics of
JPs under a linear shear flow.

Under a Taylor-Green flow, the amphiphilic JPs are the most interesting,
exhibiting different vesicle shapes depending on the ratio of the size
of the TG flow to the vesicle size. These results show that the shape of
the vesicle, whether square or polygonal, can be controlled by adjusting
the size of the cell in the TG flow. On the other hand, the assembly of
multi-lamella JPs and the striated JPs behave more like a rigid body
with connected subdomains, and the number of subdomains increases with
increasing strength of the TG flow. Overall, the multilamellar (BC (ii))
JP assembly behaves as a shear-thinning fluid, while the striated (BC
(iii)) JP assembly  possesses a yield stress.

The results reported here provide an inroad modeling framework for
hydrodynamics of active colloids~\cite{Meredithetal2022,
McGlassonBradley2021, Vutukuri2020, Mallory2017}. The present study also
helps us understand the rheology of JP oligomers that may be realized in
the experiments. We are extending this study to three dimensional
systems with more realistic features such as size distributions of JPs
and thermal fluctuations~\cite{kohl-cor-che-vee22}. From a numerical
perspective, it is straightforward to include random perturbations in
the particle shape and boundary condition that mimic interfacial
properties found in lab conditions~\cite{Bradley2016, Bradley2017,
Zarzaretal2015, doi:10.1021/la503455h}.

\section{Appendix}
\label{sec:appendix}
The boundary integral calculation
of \eqref{eq:free_energy2} relies
on the following identity 
%
\begin{equation}
\label{eq:normal_deriv}
\nabla u(\xx) \cdot \nnu_\xx=
%\frac{\partial}{\partial \nnu_\xx} \DD[\sigma](\xx) =
-\frac{1}{\rho^2} {\bf t}_\xx\cdot \SSS[\sigma{\bf t}](\xx)
+ \frac{\dif}{\dif s}\SSS\left[\frac{\dif \sigma}{\dif s}\right](\xx), \quad \xx \in \partial \Omega.
\end{equation}
%
Here, $\tt_\xx$ is the tangent vector and $\dif/\dif s$ is the
arclength derivative.
Substituting~\eqref{eq:normal_deriv}
into~\eqref{eq:free_energy2} for the normal derivative leads to the
single layer, $\SSS$, which is more straightforward to evaluate.
The arclength derivative is computed using
the spectrally accurate Fourier differentiation.

@@

% If you have acknowledgments, this puts in the proper section head.
\begin{acknowledgments}
%We thank Dr.~S\'ebastien Michelin and Dr.~Cecile Cottin-Bizzonne from LadHyX - Ecole Polytechnique Institut Polytechnique de Paris for organizing the focus-session ``interfacial active matter'' in the conference APS-DFD 2021 and the invitation of submitting a special collection paper. 
We thank useful conversations with E.~Corona, M.~Rachh, S.~Jiang, and
  T.~Beck.  B.Q.~acknowledges support from the Simons Foundation,
  Mathematics and Physical Sciences-Collaboration Grants for
  Mathematicians, Award No.~527139.  Y.-N.Y.~acknowledges support from
  NSF (Grants No.~DMS 1614863 and No.~DMS 195160) and Flatiron
  Institute, part of Simons Foundation.

@@

Multilamelar bilayers arise when both sides of the particle 
are hydrophobic. 
The boundary condition \eqref{eq:bcs}(ii) 
gives a particle with a hydrophobic intensity that is greater
on the $\theta_i = 0$ side than on the $\theta_i = \pi$ side.
The initial self-assembly is similar to that in case (i).
The difference arises in the long-time dynamics where the bilayers
no longer remain well-separated. 
Rather, the bilayers form layers 
as a consequence of the interfacial tension of exposed particle
heads.%~\cite{Huetal19, deMeetal21}. 
Fig~\ref{fig:relax}(c) shows the multilamellar structure with 8 layers when $N_b=60$. 
The number of folds depends on several factors, for instances, number of
particles and initial configurations.

Finally, boundary condition \eqref{eq:bcs}(iii) models 
a particle whose head surface repels the tail surface as proposed in
\cite{MaRa76, Ma77}.
The particles initially form chains with their directors perpendicular to the
length of the chain. The equilibrium structure
resembles a checkerboard pattern
where each particle coordinates
its head with the head of three other particles and its tail with the
tail of three other particles. % Fig~\ref{fig:self-assembly}(c).
Fig~\ref{fig:relax}(d) gives an example that this boundary condition results a checker board/stripes equilibrium state.

@@

Consider placing the equilibrium state of boundary condition \eqref{eq:bcs} (ii) for 60 particles in a shear flow,
\begin{equation}
\uu_\infty(\xx) = \dot\gamma ({\bf e}_y\cdot\xx){\bf e}_x,
\end{equation}
%
where $\dot\gamma$ is the shear rate. Fig~\ref{fig:shear_1} (a)--(c) show the snapshots of simulation results when $\dot\gamma=0.05$.
We measure the ratio of major and minor axes $\lambda$ by adopting an ellipse fitting using the singular value decomposition and track this aspect ratio for this multi-lamellar structure when the flow is on. 
The result is shown in panel (d) and we observe that the multi-lamellar structure will be slightly stretched and the aspect ratio $\lambda=3.5$ increases over time.
%
Notice that for this simulation, the center of mass position of the
initial configuration is not at the origin and it is slightly to the
right of the origin. We also track the aspect ratio $\lambda$ for the
case that the center of mass position is shifted to the origin. Panel
(d) also shows that the dynamics of two sets of simulations are
invariant of center of mass position in a shear flow.

@@

Similar to shear flow simulations, we then place equilibrium states of all boundary conditions \eqref{eq:bcs} in a rotating Taylor-Green Flow with a length scale $l$,
\begin{equation}
\uu_\infty(\xx) = V_0 \left(-\cos(l{\bf e}_x\cdot\xx)\sin(l{\bf e}_y\cdot\xx){\bf e}_x+\sin(l{\bf e}_x\cdot\xx)\cos(l{\bf e}_y\cdot\xx){\bf e}_y\right).
\end{equation}
%
The magnitude of the vortex flow is controlled by the velocity scale
$V_0$.  Similar to the setup in the shear flow cases, we shift the
center of mass position to the origin then turn on the Taylor-Green flow
with $V_0=0.1$ and tune the sides of zero-velocity box to $\pi$, i.e.,
$l=0.5$, Figure~\ref{fig:TG_1}(a)--(c) show three frames of simulation
results and we track the aspect ratio $\lambda$ described in the
previous section which is provided in panel (d).  The stiffness of the
multi-lamellar structure is able to compete with the flow therefore the
structure does not deform to any states other than multi-lamellar shape.
We also found that the flow strength coefficient above $V_0=0.1$ will
result to clear group separations during the simulations.  Comparing
with the case when $V_0=0.2$, we provide a frame in
Figure~\ref{fig:TG_2}(a) and the energy comparison is shown in panel
(b). Numerical results validate the phenomenon that a higher
Taylor-Green vortex flow generates large structure separations with
higher total energy.

@@

Here are demonstrations of having JP vesicle and multi-lamellar
structure in a narrow channel under a shear flow.  Simulation results
and snapshots are shown in Figures~\ref{fig:confined_vesicle} and
Figure~\ref{fig:confined_lamellar} where the JP vesicle has $N_b=25$
particles and JP lamellar has $N_b=40$ particles, respectively. Both
starting configurations are pre-relaxed and the channel width in the
narrow part is about 5 length units in the vesicle case and 4 length
units in the lamellar case.

@@

\begin{figure}
  \begin{center}
  \includegraphics[width=0.4\textwidth]{MultConf_100.png}
  \includegraphics[width=0.4\textwidth]{MultConf_600.png}\\
  \includegraphics[width=0.4\textwidth]{MultConf_1200.png}
  \includegraphics[width=0.4\textwidth]{MultConf_2400.png}
  \end{center}
  \vspace{-20pt}  
  \caption{\label{fig:confined_lamellar} A JP multi-lamellar structure confined in a narrow channel where $N_b=40$ and $\dot\gamma=0.05$.  }
\end{figure}

>>>>

\begin{abstract}
  Active colloidal systems with non-equilibrium self-organization is a
  long-standing, challenging area in material sciences and biology. To
  understand how hydrodynamic flow may be used to actively control
  self-assembly of Janus particles (JPs), we developed a model for the
  many-body hydrodynamics of amphiphilic JPs suspended in a viscous
  fluid with imposed far-field background flows (Journal of Fluid
  Mechanics, 941, 2022). In this work we alter the hydrophobic
  distribution on the JP-solvent interface to investigate the
  hydrodynamics that underlies the various morphologies and rheological
  properties of the JP assembly in the suspension. We find JPs assemble
  into unilamellar, multilamellar, and striated structures. To introduce
  dynamics, we include a planar linear shear flow and a steady
  Taylor-Green mixing flow, and measure the collective dynamics of JP
  particles in terms of their (a) free energy from the hydrophobic
  interactions between the JPs, (b) order parameter for the ordering of
  JPs in terms of alignment of their directors, and (c) strain parameter
  that captures the deformation in the assembly. We characterize the
  effective material properties of the JP structures and find that the
  unilamellar structures increases orientation order under shear flow,
  the multilamellar structure behaves as a shear thinning fluid, and the
  striated structure possesses a yield stress. These numerical results
  provide insights into dynamic control of non-equilibrium active
  biological systems with similar self-organization.
\end{abstract}

@@

The shear stress-strain rate data are excellently fit by power laws.
Figures~\ref{fig:rheologysummary}(a)--(c) show the data for Type I, Type
II and Type III JPs, respectively, under both shear flow (squares) and
TG flow (triangles). The bilayer phases consisting of Type I JPs are
basically Newtonian with viscosity $0.014$ MPa\;$\upmu$s $= 14$ mPa\;s,
about 14 times the viscosity of water at room temperature
(Figures~\ref{fig:rheologysummary}(a), dashed curve). The vesicle
bilayer phase is slightly shear thickening
(Figure~\ref{fig:rheologysummary}(a), solid curve). The multilamellar
phase consisting of Type II JPs is somewhat shear thinning, but its
viscosity is larger than for Type I. Finally, the striated phase is
strongly shear-thinning. This further explains the rigid-like behavior
of the striated phase found in the strain, shear-rate relationships from
Figures~\ref{fig:s2esummary}(a) and (b) (triangles).
\todo[inline]{RBC flow references}

@@

In this work, we employ the newly developed JP model using
BIEs~\cite{Fu20, Fu2022_JFM} and tune the boundary conditions with
energy normalization to study the collective dynamics of amphiphilic
(Type I), biased hydrophobic (Type II), and bipolar (Type III) JPs under
various flowing conditions (quiescent flow, linear shear flow, and
Taylor-Green flow). Three quantities are computed to characterize the
dynamics of the collective configurations of JPs: free energy $F$,
strain parameter $E$, and scalar order parameter $S_2$. 
%We quantify how
%the collective hydrodynamics of JPs is affected by an external flow.
Under a given flow, we use these three measures of deformations to
quantify the differences in the collective dynamics between the three
types of JPs. These results, summarized below, provide general insight
into the dynamic control of active particles in a viscous suspension.

%Within the framework of boundary integral equations, we derive a
%convenient expression for computing the free energy using the
%single-layer potential (see equation~\eqref{eq:normal_deriv}).
In a quiescent flow, the free energy profiles demonstrate that the
relaxation process for particles confined in a certain size of box is
independent of the number of particles. However, we find that the final
configurations does depend on the initial distribution of particle
directors $\dd_i$. Therefore, multiple patterns or local energy minimum
states may appear depending on the initial setup.

Under a relatively weak linear shear flow, the amphiphilic JPs behave as
a unilamellar vesicle that elongates and tank-treads, with the scalar
order parameter increasing over time. The assembly of multilamellar and
striated JP structures resemble a rigid body motion with minimal
deformation. The effective viscosity of the material quantitatively
validates this result. High shear-rate cases provide a range of critical
shear rates where the structures break apart and undergo topological
changes. We further show that free energy, scalar order parameter, and
strain are effective measures to quantitatively capture the collective
hydrodynamics of JPs under a linear shear flow.

Under a Taylor-Green flow, the amphiphilic JPs are the most interesting,
exhibiting different vesicle shapes depending on the ratio of the size
of the TG flow to the vesicle size. These results show that the shape of
the vesicle, whether square or polygonal, can be controlled by adjusting
the size of the cell in the TG flow. On the other hand, the assembly of
multilamellar JPs and the striated JPs behave more like a rigid body
with connected subdomains, and the number of subdomains increases with
increasing strength of the TG flow. Overall, the multilamellar (Type II)
JP assembly behaves as a shear-thinning fluid, while the striated (Type
III) JP assembly possesses a yield stress.

The results reported here provide an inroad modeling framework for
hydrodynamics of active colloids~\cite{Meredithetal2022,
McGlassonBradley2021, Vutukuri2020, Mallory2017}. The present study also
helps to understand the rheology of JP oligomers that may be realized in
experiments. We are extending this study to three dimensional systems
with more realistic features such as size distributions of JPs and
thermal fluctuations~\cite{kohl-cor-che-vee22}. From a numerical
perspective, it is straightforward to include random perturbations in
the particle shape and boundary condition that mimic interfacial
properties found in lab conditions~\cite{Bradley2016, Bradley2017,
Zarzaretal2015, doi:10.1021/la503455h}.

\section{Appendix}
\label{sec:appendix}
The boundary integral calculation of~\eqref{eq:free_energy2} relies on
the following identity 
%
\begin{equation}
\label{eq:normal_deriv}
\nabla u(\xx) \cdot \nnu_\xx=
%\frac{\partial}{\partial \nnu_\xx} \DD[\sigma](\xx) =
-\frac{1}{\rho^2} {\bf t}_\xx\cdot \SSS[\sigma{\bf t}](\xx)
+ \frac{\dif}{\dif s}\SSS\left[\frac{\dif \sigma}{\dif s}\right](\xx), \quad \xx \in \partial \Omega.
\end{equation}
%
Here, $\tt_\xx$ is the tangent vector and $\dif/\dif s$ is the arclength
derivative. Substituting~\eqref{eq:normal_deriv}
into~\eqref{eq:free_energy2} for the normal derivative leads to the
single layer, $\SSS$, which is more straightforward to evaluate. The
arclength derivative is computed using the spectrally accurate Fourier
differentiation.

@@

% If you have acknowledgments, this puts in the proper section head.
\begin{acknowledgments}
We thank useful conversations with E.~Corona, M.~Rachh, S.~Jiang, and
  T.~Beck. B.Q.~acknowledges support from the Simons Foundation,
  Mathematics and Physical Sciences-Collaboration Grants for
  Mathematicians, Award No.~527139. Y.-N.Y.~acknowledges support from
  NSF (Grants No.~DMS 1614863 and No.~DMS 195160) and Flatiron
  Institute, part of Simons Foundation.

@@

Multilamelar bilayers arise when both sides of the particle 
are hydrophobic. 
The boundary condition \eqref{eq:bcs}(ii) 
gives a particle with a hydrophobic intensity that is greater
on the $\theta_i = 0$ side than on the $\theta_i = \pi$ side.
The initial self-assembly is similar to that in case (i).
The difference arises in the long-time dynamics where the bilayers
no longer remain well-separated. 
Rather, the bilayers form layers 
as a consequence of the interfacial tension of exposed particle
heads.%~\cite{Huetal19, deMeetal21}. 
Figure~\ref{fig:relax}(c) shows the multilamellar structure with 8
layers when $N_b=60$.  The number of folds depends on several factors,
for instances, number of particles and initial configurations.

Finally, boundary condition \eqref{eq:bcs}(iii) models a particle whose
head surface repels the tail surface as proposed in \cite{MaRa76, Ma77}.
The particles initially form chains with their directors perpendicular
to the length of the chain. The equilibrium structure resembles a
checkerboard pattern where each particle coordinates its head with the
head of three other particles and its tail with the tail of three other
particles. Figure~\ref{fig:relax}(d) gives an example that this boundary
condition results a checker board/stripes equilibrium state.

@@

Consider placing the equilibrium state of boundary condition \eqref{eq:bcs} (ii) for 60 particles in a shear flow,
\begin{equation}
\uu_\infty(\xx) = \dot\gamma ({\bf e}_y\cdot\xx){\bf e}_x,
\end{equation}
%
where $\dot\gamma$ is the shear rate. Fig~\ref{fig:shear_1} (a)--(c) show the snapshots of simulation results when $\dot\gamma=0.05$.
We measure the ratio of major and minor axes $\lambda$ by adopting an ellipse fitting using the singular value decomposition and track this aspect ratio for this multilamellar structure when the flow is on. 
The result is shown in panel (d) and we observe that the multilamellar structure will be slightly stretched and the aspect ratio $\lambda=3.5$ increases over time.
%
Notice that for this simulation, the center of mass position of the
initial configuration is not at the origin and it is slightly to the
right of the origin. We also track the aspect ratio $\lambda$ for the
case that the center of mass position is shifted to the origin. Panel
(d) also shows that the dynamics of two sets of simulations are
invariant of center of mass position in a shear flow.

@@

Similar to shear flow simulations, we then place equilibrium states of all boundary conditions \eqref{eq:bcs} in a rotating Taylor-Green Flow with a length scale $l$,
\begin{equation}
\uu_\infty(\xx) = V_0 \left(-\cos(l{\bf e}_x\cdot\xx)\sin(l{\bf e}_y\cdot\xx){\bf e}_x+\sin(l{\bf e}_x\cdot\xx)\cos(l{\bf e}_y\cdot\xx){\bf e}_y\right).
\end{equation}
%
The magnitude of the vortex flow is controlled by the velocity scale
$V_0$.  Similar to the setup in the shear flow cases, we shift the
center of mass position to the origin then turn on the Taylor-Green flow
with $V_0=0.1$ and tune the sides of zero-velocity box to $\pi$, i.e.,
$l=0.5$, Figure~\ref{fig:TG_1}(a)--(c) show three frames of simulation
results and we track the aspect ratio $\lambda$ described in the
previous section which is provided in panel (d).  The stiffness of the
multilamellar structure is able to compete with the flow therefore the
structure does not deform to any states other than multilamellar shape.
We also found that the flow strength coefficient above $V_0=0.1$ will
result to clear group separations during the simulations.  Comparing
with the case when $V_0=0.2$, we provide a frame in
Figure~\ref{fig:TG_2}(a) and the energy comparison is shown in panel
(b). Numerical results validate the phenomenon that a higher
Taylor-Green vortex flow generates large structure separations with
higher total energy.

@@

Here are demonstrations of having JP vesicle and multilamellar
structure in a narrow channel under a shear flow. Simulation results
and snapshots are shown in Figures~\ref{fig:confined_vesicle} and
Figure~\ref{fig:confined_lamellar} where the JP vesicle has $N_b=25$
particles and JP lamellar has $N_b=40$ particles, respectively. Both
starting configurations are pre-relaxed and the channel width in the
narrow part is about 5 length units in the vesicle case and 4 length
units in the lamellar case.

@@

\begin{figure}
  \begin{center}
  \includegraphics[width=0.4\textwidth]{MultConf_100.png}
  \includegraphics[width=0.4\textwidth]{MultConf_600.png}\\
  \includegraphics[width=0.4\textwidth]{MultConf_1200.png}
  \includegraphics[width=0.4\textwidth]{MultConf_2400.png}
  \end{center}
  \vspace{-20pt}  
  \caption{\label{fig:confined_lamellar} A JP multilamellar structure
  confined in a narrow channel where $N_b=40$ and $\dot\gamma=0.05$.}
\end{figure}
